Reformulating phylogenetic mixed models to improve flexibility and speed

### 2019-12-04 phyloglmm_ms.tex
minor change and citations

@@

Classic phylogenetic regressions usually allow the response (trait or distribution) to evolve according to the phylogenetic relationship across species, but the effects of the predictor variables may evolve according to the phylogenetic relationship across species as well.
Suppose we have examined a collection of species that came from two groups, and wish to know whether their brain size (Y) is proportional to their body size (X) (Felsenstein's (1985) example using a mixed-effect model. 
Standard phylogenetic regressions allow for phylogenetic correlations in the intercept of the relationship between body and brain size. 
However, species within taxonomic groups with similar body sizes may vary in overall brain size, or taxonomic groups may vary in the relationship between brain and body size. 
%\bmb{maybe we can focus on random slopes here? Need to point out that example studies below are \emph{exceptions} to the general assumption of random intercepts only (how did they manage to do their inference?)}
Several recent studies have incorporated phylogenetic variation in different ways and looked at species response to phylogenetic variation with changes in environmental factors.
For example, \cite{nowakowski2018phylogenetic} considered phylogenetically correlated slopes in response to habitat conversion when studying the abundance of amphibian species using a Bayesian phylogenetic GLMM, while \cite{li2017canfun} considered phylogenetically correlated species nested within sites when modeling plant abundance via a phylogenetic GLMM approach. 
The tools available for extending phylogenetic relationships to predictor variables (or predictor level variation) in a standard frequentist framework are relatively inflexible; thus, many biologist needing to fit random-slopes model have usually turned to more flexible Bayesian approaches, despite their additional computational burden \citep{hadfield2010mcmc, burkner2018brms}.
Table \ref{table:model} summarizes types of platforms, data constraints, and provides model complexities for phylogenetic comparative analysis.

@@

In this paper, we will propose an alternative formulation of phylogenetic regression in a generalized linear mixed modeling framework that is mathematically equivalent to previous approaches, but more flexible.
In particular, our new formulation can be implemented in any framework that allows for random effects (such as random intercepts, slopes, and interactions), without the need to implement special correlation structures, by incorporating phylogenetic structures as part of the mean model. 
%% can be incorporated relatively easily into any mixed-model platform that allows new random-effects design matrices to be specified, and can be easily extended along with the features of mixed model such as multiple observational designs, phylogenetic random-slopes models, and multiple (nested or crossed) random effects.
We will compare our technique coded in R packages \pkg{lme4} and \pkg{glmmTMB} with existing R packages (i.e. \pkg{nlme} \citep{pinheiro2014r}, \pkg{phylolm} \citep{ho2014phylolm}, \pkg{pez} \citep{pearse2015pez} \pkg{phyr} (Li et al, unpublished), \pkg{MCMCglmm} \citep{hadfield2010mcmc}, and \pkg{brms} \citep{burkner2018brms}), fitting models to data from simulated model that incorporates phylogenetic variation from both predictors and tips/species as well as residual variation.
% We expect that our approach will be fast as well as flexible, because \pkg{lme4} and \pkg{glmmTMB} use efficient approaches. 
%% TODO LAST LINE SEE PAGE 5

>>>>

Classic phylogenetic regressions usually allow the response (trait or distribution) to evolve according to the phylogenetic relationship across species, but the effects of the predictor variables may evolve according to the phylogenetic relationship across species as well. 
For example, suppose we have examined a collection of species that came from two groups, and wish to know whether their brain size (Y) is proportional to their body size (X) (Felsenstein's (1985) example using a mixed-effect model. 
The standard phylogenetic mixed-effect model allows for phylogenetic correlations in the intercept of the relationship between body and brain size. 
However, species within taxonomic groups with similar body sizes may vary in brain size (random intercepts), or taxonomic groups may vary in the relationship between brain and body size (random slopes). 
Several recent studies have incorporated some form of phylogenetic variation using random effects in addition to the standard phylogenetic correlations in the intercepts and looked at species response to phylogenetic variation with changes in environmental factors.
For example, \cite{nowakowski2018phylogenetic} considered phylogenetically correlated slopes in response to habitat conversion when studying the abundance of amphibian species using a Bayesian phylogenetic GLMM, while \cite{li2017canfun} considered phylogenetically correlated species nested within sites when modeling plant abundance via a phylogenetic GLMM approach. 
The tools available for extending phylogenetic relationships to predictor variables (or predictor level variation) in a standard frequentist framework are relatively inflexible; thus, many biologist needing to fit random-slopes model have usually turned to more flexible Bayesian approaches, despite their additional computational burden \citep{hadfield2010mcmc, burkner2018brms}.
Table \ref{table:model} summarizes types of platforms, data constraints, and provides model complexities for phylogenetic comparative analysis.

@@

@@

In this paper, we will propose an alternative formulation of phylogenetic regression in a generalized linear mixed modeling framework that is mathematically equivalent to previous approaches, but more flexible.
In particular, our new formulation can be implemented in any framework that allows for random effects (such as random intercepts, slopes, and interactions), without the need to implement special correlation structures, by incorporating phylogenetic structures as part of the mean model \citep{hefley2017basis}.
%% can be incorporated relatively easily into any mixed-model platform that allows new random-effects design matrices to be specified, and can be easily extended along with the features of mixed model such as multiple observational designs, phylogenetic random-slopes models, and multiple (nested or crossed) random effects.
We will compare our technique coded in R packages \pkg{lme4} and \pkg{glmmTMB} with existing R packages (i.e. \pkg{nlme} \citep{pinheiro2014r}, \pkg{phylolm} \citep{ho2014phylolm}, \pkg{pez} \citep{pearse2015pez} \pkg{phyr} \citep{phyr}, \pkg{MCMCglmm} \citep{hadfield2010mcmc}, and \pkg{brms} \citep{burkner2018brms}), fitting models to data from simulated model that incorporates phylogenetic variation from both predictors and tips/species as well as residual variation.
% We expect that our approach will be fast as well as flexible, because \pkg{lme4} and \pkg{glmmTMB} use efficient approaches. 
%% TODO LAST LINE SEE PAGE 5
